---
title: "hw6"
author: "Senna"
date: "2024-11-30"
output: github_document
---

Necessary packages are loaded.
```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(broom)
library(ggplot2)
library(purrr)
library(modelr)
library(rsample)

set.seed(123)
```



## Problem 1
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

# estimates
# plot the distribution of estimates, describe these in words.
# identify the 2.5% and 97.5% quantiles to provide a 95% CI for each estimate. 

```{r}
boot_weather = weather_df |> 
  bootstrap(n = 5000) |> 
  mutate(
    models = map(strap, \(df) lm(tmax ~ tmin, data = df)),
    glance_results = map(models, broom::glance),
    tidy_results = map(models, broom::tidy)
  ) |> 
  select(-strap, -models) |> 
  unnest(cols = c(glance_results, tidy_results), names_sep = "_") 


boot_weather|> 
  summarize(
    r_squared_mean = mean(glance_results_r.squared, na.rm = TRUE),
    r_squared_lower_ci = quantile(glance_results_r.squared, 0.025, na.rm = TRUE),
    r_squared_upper_ci = quantile(glance_results_r.squared, 0.975, na.rm = TRUE),
    log_product = ifelse(
      all(c("(Intercept)", "tmin") %in% tidy_results_term),
      tidy_results_estimate[tidy_results_term == "(Intercept)"] * 
      tidy_results_estimate[tidy_results_term == "tmin"],
      NA_real_
    ),
    log_product_mean = mean(log(log_product), na.rm = TRUE),
    log_product_lower_ci = quantile(log(log_product), 0.025, na.rm = TRUE),
    log_product_upper_ci = quantile(log(log_product), 0.975, na.rm = TRUE)
  )

```


```{r}
boot_results = boot_weather|>
  filter(tidy_results_term %in% c("(Intercept)", "tmin")) |> 
  group_by(glance_results_r.squared) |> 
  summarize(
    r_squared = glance_results_r.squared,
    log_product = log(
      tidy_results_estimate[tidy_results_term == "(Intercept)"] *
      tidy_results_estimate[tidy_results_term == "tmin"]
    )
  )
```
###### r2 plot
```{r}
ggplot(boot_results, aes(x = r_squared)) +
  geom_density (color ='red', alpha = 0.5)+
  theme_minimal() +
  labs(
    title = "Distribution of R^2 from Bootstrapped Samples",
    x = expression(R^2),
    y = "Frequency"
  )
```

###### log product plot
```{r}
ggplot(boot_results, aes(x = log_product)) +
  geom_density (color ='blue', alpha = 0.5)+
  theme_minimal() +
  labs(
    title = "Distribution of log(β^0 ∗ β^1) from Bootstrapped Samples",
    x = expression(log(beta^0 * beta^1)),
    y = "Frequency"
  )
```

The log product follows a normal distribution. We had n=5000 for bootstrapping, which is a large sample size. This invokes the Central Limit Theorem.

r^2 plot shows a left skew. r^2 values are bounded between 0 and 1, and in our case, the estimate is close to 1. 

## Problem 2

```{r}
homicide_df = read.csv("./data/homicide-data.csv")|>
  mutate(
    city_state = paste(city, state, sep=","),
    solved = if_else(disposition == "Closed by arrest",1,0),
    victim_age = as.numeric(victim_age),
    victim_sex = as.factor(victim_sex),
    victim_race = as.factor(victim_race),
    solved = as.factor(solved)
  )

filtered = homicide_df|>
  filter(
    !city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"),
    victim_race %in% c("White", "Black"),
    !is.na(victim_age)
  )
```

```{r}
baltimore_df = filtered |>
  filter(city_state == "Baltimore,MD")

baltimore_model = glm(
  solved ~ victim_age + victim_sex + victim_race,
  family = binomial,
  data = baltimore_df
)

baltimore_results = tidy(baltimore_model, conf.int = TRUE, exp = TRUE) # true true?

sex_diff = baltimore_results|>
  filter (term == "victim_sexMale")|>
  select(estimate = estimate, conf.low, conf.high)

print(sex_diff)
```

```{r}
city_model = filtered|>
  group_by(city_state)|>
   filter(
    n_distinct(victim_sex) > 1,  
    n_distinct(victim_race) > 1   
  ) |>
  nest()|>
  mutate(
    model = map(data, ~glm(solved ~ victim_age + victim_sex + victim_race, family = binomial, data = .x)),
    results = map(model, ~tidy(.x, conf.int = TRUE, exp = TRUE))
  )|>
  unnest(results)|>
  filter(term == "victim_sexMale")|>
  select(city_state, estimate = estimate, conf.low, conf.high)
```

```{r}
ordered_results = city_model|> arrange(estimate)

ggplot(ordered_results, aes(x= reorder(city_state, estimate), y = estimate))+
  geom_point()+
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high))+
  labs(
    title = "adjusted odds ratio for solved homicides",
    x = "city",
    y = "adjusted odds ratio: male vs female"
  )+
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1)  
  )
```
COMMENT ON THE PLOT

## Problem 3

```{r}
bw_df = read.csv("./data/birthweight.csv")
glimpse(bw_df)
```

```{r}
bw_df <- bw_df |>
  mutate(
    babysex = factor(babysex, labels = c("Male", "Female")),
    frace = factor(frace, levels = c(1, 2, 3, 4, 8, 9), 
                   labels = c("White", "Black", "Asian", "Puerto Rican", "Other", "Unknown")),
    mrace = factor(mrace, levels = c(1, 2, 3, 4, 8), 
                   labels = c("White", "Black", "Asian", "Puerto Rican", "Other")),
    malform = factor(malform, levels = c(0, 1), labels = c("Absent", "Present"))
  )

bw_df |>
  summarize(across(everything(), ~ sum(is.na(.))))
```
no missing values.




The factors influencing birth weight (bwt) can include:

Biological factors: bhead, blength, gaweeks, malform.
Maternal factors: ppbmi, ppwt, wtgain, momage, mheigth, smoken.
Socioeconomic factors: fincome, frace, mrace.

```{r}
# proposing a regression model

bwt_model <- lm(bwt ~ bhead + blength + gaweeks + malform + ppbmi + ppwt + wtgain + momage + mheight + smoken + fincome + frace, 
                data = bw_df)

summary(bwt_model)

```
```{r}
# model residual vs fitted values

bw_df = bw_df |>
  add_predictions(bwt_model)|>
  add_residuals(bwt_model)

ggplot(bw_df, aes(x=pred, y=resid))+
  geom_point()+
  geom_hline(yintercept=0, linetype = 'dashed')+
  labs(
    title = "residuals vs fitted values",
    x = "predicted birth weight (fitted)",
    y = "residuals"
  )
```

```{r}
# length and gestational age
model2 <- lm(bwt ~ blength + gaweeks, data = bw_df)

# head circumference, length, sex, all interactions
model3 <- lm(bwt ~ bhead + blength + babysex + bhead*blength + bhead*babysex + blength*babysex + bhead * blength * babysex, data = bw_df)
```




```{r}
cv_df = 
  crossv_mc(bw_df, 100)

cv_df =cv_df |> 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))
```

```{r}
cv_df = cv_df |> 
  mutate(
    model1  = map(train, \(df) lm(bwt ~ bhead + blength + gaweeks + malform + ppbmi + ppwt + wtgain + momage + mheight + smoken + fincome + frace, data = df)),
    model2  = map(train, \(df) lm(bwt ~ blength + gaweeks, data = df)),
    model3  = map(train, \(df) lm(bwt ~ bhead + blength + babysex + bhead*blength + bhead*babysex + blength*babysex + bhead * blength * babysex, data = df))) |> 
  mutate(
    rmse_1 = map2_dbl(model1, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_2 = map2_dbl(model2, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_3 = map2_dbl(model3, test, \(mod, df) rmse(model = mod, data = df)))
```

```{r}
cv_df |> 
  select(starts_with("rmse")) |> 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") |> 
  mutate(model = fct_inorder(model)) |> 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```







